Comparing rolling period forecasts of Facebooks' Prophet model to an ARIMA model: A Diebold-Mariano evaluation of the FTSE/JSE Top40 Index

### 2017-11-06 template/Draft_Paper.tex
Final draft with all edits

\begin{center}
{\Huge \bf Comparing Facebook's Prophet Forecasting to an ARIMA Model: A
Diebold-Mariano Evaluation of the Top40 Index}\\
\today\\
Marvelous Mubenesha ( MBNMAR005)\\
{\tt MBNMAR005@uct.co.za}
\end{center}

\begin{abstract}
Facebook's data science team open-sourced Prophet, a package that allows
analysts to forecast a wide range of business time series at-scale.
Prophet employs intensive Bayesian modelling in two tiers. Purely
through MAP parameter estimation, and in a quasi-Bayesian form through
an automated forecast evaluation that enables analyst to improve the
model if it falls short when compared to other models. This paper assess
the predictive accuracy of Prophets' forecasts to those of an ARIMA
model that was identified using the Box-Jenkins methodology. This is
achieved through a Diebold-Mariano evaluation of the rolling period
forecast errors of returns on the FTSE/JSE Top40 Index across daily,
weekly and monthly time horizons. The results of the study suggest that
there is insufficient evidence to conclude that the forecasting models
have unequal predictive ability within the scope of the JSE/Top40 Index.
\noindent
Keywords: prophet forecasting, arima forecasting, Box-Jenkins methodology,Diebold-Mariano test,multi-step rolling window forecasts, FTSE/JSE Top40 index.
\end{abstract}
\end{titlepage}

@@

In business, including that involving the stock market, analysts require
a large number of time series forecasts to inform decision making. The
challenge with existing reliable approaches to forecasting time series
data are that the models require a considerable background in statistics
to build and adapt to prevailing market conditions when needed. Several
authors have recommended automatic forecasting, which refers to
algorithms that can automatically forecast a large number of univariate
time series. However, the forecasts generated by these methods have been
found to be brittle as well as inflexible in allowing for prior
information to be incorporated into the forecasting model (Taylor and
Letham 2017). In February of 2017, Facebook's Core Datascience team open
sourced ,Prophet, their time series forecasting tool whose main selling
point is that it enables ``forecasting-at-scale''. A term coined by the
data science team which is defined as, ``an approach that allows a large
number of analysts to forecast a large number and variety of business
time series'' (Taylor and Letham 2017). It has the added advantage of
being less `expensive' than other alternatives. In addition, it is
specifically prefarable over ARIMA models because of its non-linearity,
flexibility and ability to accomodate varying time intervals (Taylor and
Letham 2017).

The objective of this paper is to compare the rolling period forecasts
generated by Prophet (which uses automated, intensive Bayesian Modelling
to `forecast-at-scale'), to those obtained by an ARIMA model
(constructed through the Box-Jenkins methodology), through a
Diebold-Mariano evaluation of the JSE Top40 Index. The study is
particularly useful because analysts from a wide background can
automatically forecast stock price returns, and incorporate their prior
knowledge in the model , without a wholesome understanding of the
statistical intricacies involved.

The paper is organized as follows. A review of literature relating to
stock price forecasting with ARIMA models and other alternative
approaches is introduced to give the reader background information on
research relating to the subject matter. Thereafter, theory relating to
the mathematical formulation and estimation of both the ARIMA and
Prophet models is included in sections 2.2 and 2.3 respectively, to
build the readers understanding of what the models look like and how
they operate. Information relating to the data that was used in this
study is contained in section 3, after which, the methodology used to
answer the research question is described in section 4. Section 5,
contains the results obtained when the methodology was executed, whilst
section 6 includes a discussion of the results and how they can be
interpreted to answer the research question. Lastly, the conclusion in
section 7 highlights the main findings of the study and recommends
further areas of research.

@@

In this section we will briefly review literature that relates to the
comparison of different approaches to forecasting stock returns. This
will include both traditional approaches as well new methods such as
automatic forecasting, machine learning techniques and quasi-Bayesian
forms which incorporate analysts' prior information. The review of
literature is concluded with a brief introduction to the Diebold-Mariano
evaluation and an outline of its suitability for the purpose of this
study. The theoretical and mathematical formulation of an ARIMA
forecasting model is then briefly introduced. Thereafter, the
theoretical framework and mathematical formulation of Facebooks' Prophet
forecasting tool is discussed. Specific attention is drawn to Prophets
distinguising features to give the reader an understanding of the value
that this forecasting technique can potentially add to stock price
prediction. Lastly, the Diebold-Mariano Evaluation technique is
introduced as a method of formally testing superior predictive accuracy
between two forecasting models.

@@

ARIMA models have been widely used as a standard model to forecast
financial data and have been shown to yield acceptable forecasting
errors (Adebiyi, Adewumi, and Ayo 2014). Adebiyi, Adewumi, and Ayo
(2014) used the Box-Jenkins methodology to build short-term stock price
prediction models for stocks on the New York Stock Exchange(NYSE) and
the Nigerian Stock Exchange(NSE). Their results showed that the
prediction errors of the model were within acceptable bounds.

ARIMA models rely on the assumption that residuals are
heteroskedastic(they have a constant variance) and normally distributed,
however, some financial data displays heteroskedasticity making GARCH
models more applicable. This theory is backed by a study conducted by
Mutendadzamera and Mutasa (2014) who compared the ability of ARIMA and
GARCH models to forecast stock prices on the Zimbabwean Stock Exchange
(ZSE). They found that the GARCH model outperforms the ARIMA model which
suggests that the residuals are heteroskedastic. This is a case in point
for emerging market stocks. The researchers noted that poor liquidity in
the market could be a cause of the results they observed (Mutendadzamera
and Mutasa 2014).

@@

Though the ARIMA model is tractable given that it is simple,
interpretable and yields forecasts that are significantly accurate when
compared to other methods, it has its limitations. The most popular
being its inability to capture non-linear patterns in data, even after
its evolution from the standard form to more adaptable formulations
(Moreno, Pol, and Gracia 2011). Over the past two to three decades with
the evolution of computational power and statistical advancement, other
stock price prediction methods have been proposed, the most prominent
being a machine learning approach called Artificial Neural
Networks(ANNs)(Lin, Yang, and Song 2009). Also popular amongst the new
stock price forecasting approaches are hybrids of existing methods that
incorporate the benefits of different approaches.

Artificial Neural Networks (ANNs) are a multi-layered perceptron as
illustrated in figure 1.\\
\includegraphics{ANN.png}\\
\emph{Figure 1 : Illustration of an Artificial neural network} (Tanikic
and Despotovic 2012)\\
An ANN consists of a sorted triple \((K, A, \omega)\). \(K\) is a set
representing the multiple layers/levels of the network i.e.~input,
hidden and output levels. A is a set of pairs that represents index for
the estimate in the \(i^{th}\) node of the \(k^{th}\) layer of the
network. In the enlarged section of Figure 1, \(a_{i,k} \in A\). Lastly,
\(\omega\) represents a function that defines the weights
(\(\omega_{i,r}\)) of connections between neurons/nodes (Kriesel 2007).
Artificial Neural networks apply iterated optimization of model
parameters across the network in the form of weights conditional on
observed values to learn (Segaran 2007). Several studies have compared
ARIMA model forecasts to ANN and their conclusions are contradictory.

Researchers have found that the performance of either method depends on
the nature of the data and forecasting problem (Kihoro, Otieno, and
Wafula 2004). Stock price data is proposed to be nonlinear which
suggests that nonlinear approaches have the potential to produce better
forecasts than linear models. Furthermore, ANNs make no assumptions
about the distribution of the errors as compared to the linear ARIMA
model (Adebiyi, Adewumi, and Ayo 2014). Adebiyi, Adewumi, and Ayo (2014)
compared NYSE stock index forecasts of an ARIMA model to those of an ANN
and found that the forecasting accuracy of the ANN model was superior to
that of the ARIMA model. It is evident that ANN are preferred to ARIMA
models as a model free, nonlinear alternative. However, the model
construction of an ANN requires trial and error to initialize parameter
estimates and these parameters are not easily interpretable by analysts
(Moreno, Pol, and Gracia 2011). Researchers have found different ways to
guide analysts in this respect. In 1996, Wang and Leu (1996) proposed an
ARIMA-based ANN to forecast the medium-term price of the Taiwan Stock
Exchange Weighted Stock Index (TSEWSI). They used the Box-Jenkins
methodology to difference the series and then trained the data on a
neural network with initialisations that were guided by their
observations of the ARIMA model. This approach yielded forecasts with an
acceptable prediction accuracy based on residual analysis of the
out-of-sample data. Wang and Leu therefore concluded that the
ARIMA-based Neural Network outperformed a Neural Network trained using
raw stock price data (Wang and Leu 1996). Zhang and Wu (2009) used a
combination of the backpropagation algorithm from Neural Networks with
Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that
forecasts the S\&P 500 stock index by minimizing the mean square error.
Model forecasts were evaluated through simulation experiments and the
results led to the conclusion that the hybrid model produced superior
forecasts (Zhang and Wu 2009). This study further highlighted the
potential that nonlinear approaches have in forecasting stock prices. An
issue that arises with such methods is the complexity of the proposed
models which limits the flexibility of unseasoned analysts to adjust
model parameters as a way of improving forecasting accuracy. As we will
see, Prophet forecasting elegantly deals with this issue in the form of
a nonlinear, Generalized Additive Model (GAM).

\subsubsection{Quasi-Bayesian methods incorporating Analysts'
knowledge}\label{quasi-bayesian-methods-incorporating-analysts-knowledge}

The stock price forecasting approaches considered so far purely apply
technical data analysis methods to generate stock price forecasting
models. However, studies that have compared pure technical approaches to
those that incorporate the analysts knowledge in a quasi-Bayesian form
have been shown to be superior at predicting market prices (Givoly and
Lakonishok 1984 \& Guerard (1989)). The first instance of combining
time-series model forecasts and an analysts forecasts to obtain superior
forecasts of a stocks annual earnings can be dated to 1989. Guerard
(1989) used an additive model to combine consensus security analyst
forecasts from the S\&P Annual Earnings forecaster and annual earnings
forecasts generated by an ARIMA model with a constant. The combined
model that was estimated using ordinary least squares reduced the mean
square error of the time series and analysts forecasts from 1.28 and
1.27, respectively to 1.04. These results suggest that analysts can
substantially reduce forecasting errors by combining both approaches
(Guerard 1989). Zahedi and Rounaghi (2015) applied the same overarching
principle when they used principle component analysis to determine an
appropriate input variable, which they then used to train an ANN in an
attempt to predict stock prices on the Tehran Stock Exchange. Their
model yielded acceptable forecasting errors and performed better than
the pure ANN which further reiterates the potential of incorporating an
analysts knowledge to improve stock price forecasts, in both linear and
nonlinear models.

\subsubsection{Diebold-Mariano
Evaluation}\label{diebold-mariano-evaluation}

As is evident from the previous sections, the modelling approach and
assumptions of Prophet differ to those of an ARIMA model. Since the two
methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as
a means of testing forecasting accuracy. This is because the DM Test is
a model free test of forecasting accuracy. It is applicable to a wide
range of situations from multi-period forecasts, non-Gaussian forecast
residuals, non-quadratic loss functions and even serially correlated
data(Mariano 2000).

\subsection{ARIMA Model Formulation}\label{arima-model-formulation}

The ARIMA model is a generalisation of Autoregressive Moving Average
(ARMA) models which combine autoregression and moving average features
(Box and Jenkins 1970). In the context of returns, the ARIMA formulation
for the return at time index \(t\), \(r_t\), is given by,
\[  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} \]\\
where:\\
\(\bullet \phi_i\) and \(\psi_j\) are the parameters to be estimated;\\
\(\bullet r_{t-k}\) is the \(k^{th}\) lagged return;\\
\(\bullet \epsilon_t\) is the error term at time \(t\) which is assumed
to be white noise and\\
\(\bullet\epsilon_{t-k}\) is the \(k^{th}\) lagged error.\\
\(\bullet\)p is the order of the autoregressive component (dependenceon
history of \(r_t\)) \(\bullet\)q is the order of the moving average
component (dependenceon history of \(\epsilon_t\))

@@

\subsection{Prophet Model Formulation and
Estimation}\label{prophet-model-formulation-and-estimation}

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine-learning approaches are that they usually
require a sufficient statistical background. Furthermore, these
approaches tend to be inflexible with respect to incorporating prior
information. Though this can be partially achieved by initializing
parameter estimates for some methods (such as ANNs) ,an analyst with no
statistical training can have difficulty interpreting and thus modifying
parameter estimates(Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes performance analysis evaluation with the interaction
of an analyst in the loop (Taylor and Letham 2017). The model is
configurable which allows an analyst to incorporate their knowledge of
the behaviour of the series into the model building process through
easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

@@

The first step of the forecasting procedure begins with the analyst
simply specifying a general model. Prophet then goes into the automated
process of estimating model parameters and doing forecast evaluation. If
forecasts are considered problematic based on evaluations of the
Simulated Historical Forecast (SHF) errors, then the package will
surface these problems. The process then transitions to the top left of
the cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst
can then use visually presented aspects of the results that raised the
issues to modify model parameters before a better model is then
estimated. The loop is repeated until forecast evaluation surfaces no
problems.

\includegraphics[width=1.00000\textwidth]{analyst in the loop.png}\\
\emph{Figure 2: Illustration of Prophets' automated forecasting
procedure with an analyst in the loop} (Taylor and Letham 2017)

The interpretability of Prophets' model parameters stems from the
mathematical formulation of the forecasting model. The formulation
posits that the series is generated by an additive, parametric function
of time \(r(t)\), defined by,
\[ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.\] The additive components
include growth, seasonality and a component that adjusts for holidays as
well as once-off events e.g.~anticipated market shocks in the instance
of financial time series.

The growth component, \(g(t)\), is modelled using a generalized form of
the logistic population growth model and is of the form;\\
\[g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}\]\\
Where:\\
\(\bullet C(t)\) represents the carrying capacity of the growth
component and can be modelled using a polynomial function of time, the
simplest being a constant or linear model.\\
\(\bullet (k+\boldsymbol{a}(t)^T\delta)\) is the growth rate factor with
the \(\boldsymbol{a}(t)^T\delta\) term enabling the forecaster to choose
where the growth rate changes( i.e change points) and\\
\(\bullet (b+\boldsymbol{a}(t)^T \gamma)\) is the adjusted offset
parameter.

The growth component of the model has two useful features that allow the
carrying capacity and growth rate of the model to vary with time. This
enables an analyst to manually define when and how the growth rate
changes at different change points (Taylor and Letham 2017). This
feature is particularly useful for forecasting stock price data since
analysts are able to incorporate market movements that lead to the
growth rate either decreasing, increasing or becoming constant.\\
The seasonality component, \(s(t)\) with periodicity \(P\) is given by,
\[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]

Yearly and weekly seasonality is modelled using the standard Fourier
series. Analysts can thus use prior knowledge to account for the effects
of periodicities they have observed seasonally which is similar to the
periodicity of Seasonal ARIMA models.

@@

Prophet translates model parameter estimation into a curve-fitting
exercise by employing Maximum Apriori Posterior (MAP) estimation, a
Bayesian approach to estimating optimal parameter values for the
forecasting model. MAP estimation finds the maximum posterior estimates
for the parameters by using the likelihood function and an apriori
distribution to evaluate a posterior distribution which is then
optimized to return point estimates of the parameters (Taylor and Letham
2017). Thereafter, the package automatically evaluates the forecasts by
comparing the Simulated Historical Forecast (SHF) errors of baseline
forecasting methods (including ARIMA forecasts) to those of Prophet. SHF
errors are forecast errors generated from random points in the history
of the time series model. In Prophet, SHF errors are compared visually
to allow the analysts to easily adjust the interpretable parameters
before automatically estimating the model and evaluating forecasts.

>>>>

\begin{center}
{\Huge \bf Comparing rolling period forecasts of Facebooks' Prophet model to an
ARIMA model: A Diebold-Mariano evaluation of the FTSE/JSE Top40 Index}\\
\today\\
Marvelous Mubenesha ( MBNMAR005)\\
{\tt MBNMAR005@myuct.co.za}
\end{center}

\begin{abstract}
Facebook's data science team open-sourced Prophet, a package that allows
analysts to forecast a wide range of business time series at-scale.
Prophet employs intensive Bayesian modelling in two tiers. Purely
through MAP parameter estimation, and in a quasi-Bayesian form through
an automated forecast evaluation that enables analyst to improve a
forecasting model if it underperforms when compared to other traditional
models. This paper assesses the predictive accuracy of Prophets'
forecasts to those of an ARIMA model that was identified using the
Box-Jenkins methodology. This is achieved through a Diebold-Mariano
evaluation of the rolling period forecast errors of returns on the
FTSE/JSE Top40 Index across daily, weekly and monthly time horizons. The
results of the study suggest that there is insufficient evidence to
conclude that the forecasting models have unequal predictive ability
within the scope of the FTSE/JSE Top40 Index.
\noindent
Keywords: prophet forecasting, arima forecasting, Box-Jenkins methodology,Diebold-Mariano test,multi-step rolling window forecasts, FTSE/JSE Top40 index.
\end{abstract}
\end{titlepage}

@@

In business, including that involving the stock market, analysts require
a large number of time series forecasts to inform decision making. The
challenge with existing reliable approaches to forecasting time series
data are that the models require a considerable background in statistics
to build and adapt to prevailing market conditions when needed. Several
authors have recommended automatic forecasting, which refers to
algorithms that can automatically forecast a large number of univariate
time series. However, the forecasts generated by these methods have been
found to be brittle as well as inflexible in allowing for prior
information to be incorporated into the forecasting model (Taylor and
Letham 2017). In February of 2017, Facebook's Core Datascience team open
sourced Prophet, a time series forecasting tool with unique features
that enable ``forecasting-at-scale''. Forecasting-at-scale is a term
coined by the data science team which they define as, ``an approach that
allows a large number of analysts to forecast a large number and variety
of business time series'' (Taylor and Letham 2017). It has the added
advantage of being less `expensive' than other alternatives. In
addition, it is specifically preferable over ARIMA models because of its
non-linearity, flexibility and ability to accomodate varying time
intervals (Taylor and Letham 2017).

The objective of this paper is to compare the rolling period forecasts
generated by Prophet (which uses automated, intensive Bayesian Modelling
to `forecast-at-scale'), to those obtained from an ARIMA model
(constructed through the Box-Jenkins methodology), through a
Diebold-Mariano evaluation of the FTSE/JSE Top40 Index. The study is
particularly useful because analysts from a wide background can reliably
forecast stock price returns automatically, and incorporate their prior
knowledge in the forecasting model, without a wholesome understanding of
the statistical intricacies involved. It is hoped that this could create
opportunities for automated invesment mechanisms.

The paper is organized as follows. A review of literature relating to
stock price forecasting with ARIMA models and other alternative
approaches is introduced in section 2.1 to give the reader background
information on existing research relating to the subject matter.
Thereafter, theory relating to the mathematical formulation and
estimation of an ARIMA and Prophet model is included in sections 2.2 and
2.3 respectively. Subsequently, in section 2.4, the theoretical
motivation and mathematical formulation of the Diebold-Mariano test is
introduced as a means of comparing the predictive accuracy of two
prediction models. Information relating to the data that was used in
this study is contained in section 3, after which, the methodology used
to answer the research question is described in section 4. Section 5,
contains the results obtained when the methodology was executed, whilst
section 6 includes a discussion of the results. The discussion includes
a response to the research question as well as some useful insights from
the results. Furthermore, it highlights some recommendations for further
areas of research based on the findings of the paper. Lastly, the
conclusion in section 7 summarizes the results of the study relative to
the main objective of the paper.

@@

In this section we will briefly review literature that relates to the
comparison of different approaches to forecasting stock returns. This
will include both traditional approaches as well as relatively new
methods such as automatic forecasting, machine learning techniques and
quasi-Bayesian forms which incorporate analysts' prior information. The
theoretical and mathematical formulation of an ARIMA forecasting model
is then briefly introduced. Thereafter, the theoretical framework and
mathematical formulation of Facebooks' Prophet forecasting tool is
discussed. Specific attention is drawn to Prophets distinguising
features to give the reader an understanding of the value that this
forecasting technique can potentially add to stock price prediction. The
background information is concluded with an introduction to the
Diebold-Mariano Evaluation as a suitable framework to formally compare
the predictive accuracy of two forecasting models.

@@

ARIMA models have been widely used as a standard model to forecast
financial data and have been shown to yield forecasting errors that lie
within acceptable bounds (Adebiyi, Adewumi, and Ayo 2014). Adebiyi,
Adewumi, and Ayo (2014) used the Box-Jenkins methodology to build
short-term stock price prediction models for stocks on the New York
Stock Exchange(NYSE) and the Nigerian Stock Exchange(NSE). Their results
showed that the predictive power of the ARIMA model was satisfactory
based on forecast error diagnostics.

ARIMA models rely on the assumption that residuals are
homoskedastic(they have a constant variance) and normally distributed.
However, the residual errors of some financial data display
heteroskedasticity making GARCH models more applicable. This theory is
backed by a study conducted by Mutendadzamera and Mutasa (2014) who
compared the ability of ARIMA and GARCH models to forecast stock prices
on the Zimbabwean Stock Exchange (ZSE). They found that the GARCH model
outperforms the ARIMA model which suggests that incorporating
heteroskedasticity of the residual errors improves the forecasting
ability of the model. This is a case in point for emerging market
stocks. The researchers noted that poor liquidity in the market could be
a cause of the results they observed (Mutendadzamera and Mutasa 2014).

@@

Though the ARIMA model is tractable given that it is simple,
interpretable and yields forecasts that are significantly accurate when
compared to other methods, it has its limitations. The most popular
being its inability to capture non-linear patterns in data, even after
its evolution from the standard form to more adaptable formulations
(Moreno, Pol, and Gracia 2011). Over the past two to three decades with
the evolution of computational power and statistical advancement, other
stock price prediction methods have been proposed, the most prominent
being a machine learning approach called Artificial Neural
Networks(ANNs) (Lin, Yang, and Song 2009). Also popular amongst the new
stock price forecasting approaches are hybrids of existing methods that
incorporate the benefits of different approaches.

An Artificial Neural Networks (ANNs) is a multi-layered perceptron with
nodes that simulate the action of neurons as illustrated in figure 1.\\
\includegraphics{ANN.png}\\
\emph{Figure 1 : Illustration of an Artificial neural network} (Tanikic
and Despotovic 2012)\\
Formally, an ANN consists of a sorted triple \((K, A, \omega)\). \(K\)
is a set representing the multiple levels of the network i.e.~input,
hidden and output levels. A is a set of pairs that represents the index
for the estimated parameter at the \(i^{th}\) node of the \(k^{th}\)
level in the network. This is illustrated by the curved down arrow in
Figure 1 which points to an enlarged segment of the \(i^{th}\) node in
the \(k^{th}\) level i.e. \(a_{i,k} \in A\). Lastly, \(\omega\)
represents a function that defines the weights (\(\omega_{i,r}\)) of
connections between nodes that interact across adjacent levels (Kriesel
2007). Artificial Neural networks apply iterated optimization of model
parameters across the network in the form of weights conditional on
observed values to learn (Segaran 2007). Several studies have compared
ARIMA model forecasts to ANN and their conclusions are contradictory.

Researchers have found that the performance of either method depends on
the nature of the data and forecasting problem (Kihoro, Otieno, and
Wafula 2004). Stock price data is proposed to be nonlinear which
suggests that nonlinear approaches have the potential to produce better
forecasts than linear models. Furthermore, ANNs make no assumptions
about the distribution of the errors as compared to the linear ARIMA
model (Adebiyi, Adewumi, and Ayo 2014). Adebiyi, Adewumi, and Ayo (2014)
compared NYSE stock index forecasts of an ARIMA model to those of an ANN
and found that the forecasting accuracy of the ANN model was superior to
that of the ARIMA model. It is evident that ANN are preferred to ARIMA
models as a model free, nonlinear alternative. However, the model
construction of an ANN requires trial and error to initialize parameter
estimates and these parameters are not easily interpretable by analysts
(Moreno, Pol, and Gracia 2011). Researchers have found different ways to
guide analysts in this respect. In 1996, Wang and Leu (1996) proposed an
ARIMA-based ANN to forecast the medium-term price of the Taiwan Stock
Exchange Weighted Stock Index (TSEWSI). They used the Box-Jenkins
methodology to difference the series and then trained the data on a
neural network with initialisations that were guided by their
observations of the ARIMA model. Their hybrid model yielded forecasts
that were comparably superior to the individual models used in isolation
based on an analysis of the out-of-sample residual error diagnostics.
Wang and Leu (1996) therefore concluded that the ARIMA-based Neural
Network outperformed a Neural Network trained using raw stock price data
(Wang and Leu 1996). Zhang and Wu (2009) used a combination of the
backpropagation algorithm from Neural Networks with Improved Bacterial
Chemotaxis Optimization (IBCO) to build a model that forecasts the S\&P
500 stock index by minimizing the mean square error. Model forecasts
were evaluated through simulation experiments and the results led to the
conclusion that the hybrid model produced superior forecasts (Zhang and
Wu 2009). This study further highlighted the potential that nonlinear
approaches have in forecasting stock prices. An issue that arises with
such methods is the complexity of the proposed models which limits the
flexibility of unseasoned analysts to adjust model parameters as a way
of improving forecasting accuracy. As we will see, Prophet elegantly
deals with this issue in the form of a nonlinear, Generalized Additive
Model (GAM).

\subsubsection{Quasi-Bayesian forecasting methods that incorporate an
analysts'
knowledge}\label{quasi-bayesian-forecasting-methods-that-incorporate-an-analysts-knowledge}

The stock price forecasting approaches considered so far purely apply
technical data analysis methods to generate stock price forecasting
models. However, studies that have compared pure technical approaches to
those that incorporate an analysts knowledge in a quasi-Bayesian form
have been shown to be superior at predicting market prices (Givoly and
Lakonishok 1984 \& Guerard (1989)). The first instance of combining
time-series model forecasts and an analysts forecasts to obtain superior
forecasts of a stocks annual earnings can be dated to 1989. Guerard
(1989) used an additive model to combine consensus security analyst
forecasts from the S\&P Annual Earnings forecaster and annual earnings
forecasts generated by an ARIMA model with a nonzero mean. The combined
model that was estimated using ordinary least squares reduced the mean
square error of the time series and analysts forecasts from 1.28 and
1.27, respectively to 1.04. These results suggest that analysts can
reduce forecasting errors by combining technical methods with expertise
developed through market knowledge (Guerard 1989). Zahedi and Rounaghi
(2015) applied the same overarching principle when they used principle
component analysis to determine an appropriate input variable which they
then used to train an ANN in an attempt to predict stock prices on the
Tehran Stock Exchange. Their model yielded reduced forecasting errors
and performed better than the pure ANN. These results further reiterate
the potential of incorporating an analysts knowledge to improve stock
price forecasts, in both linear and nonlinear models.

\subsection{ARIMA model formulation}\label{arima-model-formulation}

The ARIMA model is a generalisation of Autoregressive Moving Average
(ARMA) models which combine autoregression and moving average features
(Box and Jenkins 1970). In the context of returns, the ARIMA formulation
for the return at time index \(t\), \(r_t\), is given by,
\[  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} \]where:\\
\(\bullet\) \(\phi_i\) and \(\psi_j\) are the parameters to be
estimated\\
\(\bullet\) \(r_{t-k}\) is the \(k^{th}\) lagged return\\
\(\bullet\) \(\epsilon_t\) is the error term at time \(t\) which is
assumed to be white noise\\
\(\bullet\) \(\epsilon_{t-k}\) is the \(k^{th}\) lagged error\\
\(\bullet\) \(p\) is the order of the autoregressive component
(dependence on history of \(r_t\)) and\\
\(\bullet\) \(q\) is the order of the moving average component
(dependence on history of \(\epsilon_t\)).

@@

\subsection{Prophet model formulation and
estimation}\label{prophet-model-formulation-and-estimation}

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine learning approaches discussed in the literature
review are that they usually require a sufficient statistical background
to construct and modify if necessary. Furthermore, some of these
approaches tend to be inflexible with respect to incorporating prior
information. In some instances, such as with ANN, incorporating prior
information can be partially achieved through initializing parameter
estimates. However, an analyst with limited statistical training can
have difficulty interpreting and thus modifying initial parameter
estimates (Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes an automated evaluation of forecasts with the
interaction of an analyst in the loop (Taylor and Letham 2017). The
model is configurable which allows an analyst to incorporate their
knowledge of the behaviour of the series into the model building process
through easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

@@

The first step of the forecasting procedure begins with the analyst
simply specifying a general model (i.e.~modelling). Prophet then goes
into the automated process of estimating model parameters and performing
forecast evaluation. Automated forecast evaluation is undertaken by
constructing baseline forecasts using methods such as the sample mean,
ARIMA, exponential smoothing and naive estimates (e.g.~last value or
seasonal last value). Thereafter, simulated historical forecasting (SHF)
errors are evaluated by comparing forecasts of the baseline methods and
prophet to the observed historical values. SHF errors are forecast
errors generated by random points in the history of a time series model.
Prophets SHF errors that are relatively large compared to the baseline
methods are then surfaced with visual illustrations of the results. The
process then transitions from surface problems to the top left of the
cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst can
then visually inspect forecasts to identify issues and modify features
of the Prophet model before it is re-estimated in the modelling stage.
The loop is repeated until forecast evaluation surfaces no problems
(Taylor and Letham 2017).

\includegraphics[width=0.80000\textwidth]{analyst in the loop.png}\\
\emph{Figure 2: Illustration of Prophets' automated forecasting
procedure with an analyst in the loop} (Taylor and Letham 2017)

The interpretability of Prophets' model parameters stems from the
mathematical formulation of the generalised additive model. The
formulation posits that the series is generated by an additive,
parametric function of time \(r(t)\), defined by,
\[ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.\] The additive components
include growth, seasonality and a component that adjusts for holidays as
well as once-off events e.g.~anticipated market shocks in the instance
of financial time series.

Firstly, the growth component, \(g(t)\), is modelled using a generalized
form of the logistic population growth model and is of the form;\\
\[g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}\]\\
where:\\
\(\bullet\) \(C(t)\) represents the carrying capacity of the growth
component and can be modelled using a polynomial function of time, the
simplest being a constant or linear model.\\
\(\bullet\) \((k+\boldsymbol{a}(t)^T\delta)\) is the growth rate factor
with the \(\boldsymbol{a}(t)^T\delta\) term enabling the forecaster to
choose where the growth rate changes( i.e change points) and\\
\(\bullet\) \((b+\boldsymbol{a}(t)^T \gamma)\) is the adjusted offset
parameter.\\
The growth component of the model has two useful features that allow the
carrying capacity and growth rate of the model to vary with time. This
enables an analyst to manually define when and how the growth rate
changes at different change points (Taylor and Letham 2017). This
feature is particularly useful for forecasting stock price data since
analysts are able to incorporate market movements that lead to the
growth rate either decreasing, increasing or becoming constant.

Secondly, the seasonality component, \(s(t)\) with periodicity \(P\) is
given by, \[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]

Yearly and weekly seasonality is modelled using the standard Fourier
series. Analysts can thus use prior knowledge to account for the effects
of periodicities they have observed seasonally.

@@

Prophet translates model parameter estimation into a curve-fitting
exercise by employing Maximum Apriori Posterior (MAP) estimation, a
Bayesian approach to estimating optimal parameter values for the
forecasting model. MAP estimation finds the maximum posterior estimates
for the parameters by using the likelihood function and an apriori
distribution to evaluate a posterior distribution which is then
optimized to return point estimates of the parameters (Taylor and Letham
2017).

\subsection{Diebold-Mariano
evaluation}\label{diebold-mariano-evaluation}

As is evident from sections 2.2 and 2.3, the modelling approach and
assumptions of Prophet differ to those of an ARIMA model. Since the two
methods differ, the Diebold-Mariano evaluation is preferred as a means
of formally comparing the forecasting accuracy of the two models
statistically. The Diebold-Mariano evaluation is a \textbf{model free}
hypothesis test for the equal predictive accuracy of two forecasting
methods, referred to as model 1 and model 2. It is applicable to a wide
range of situations from forecast residuals that are non-Gaussian,
asymmetric loss functions, serially correlated errors and multi-period
forecasts (Mariano 2000). The Diebold-Mariano test (DM test) operates by
evaluating the forecast errors for both model 1 and model 2 (i.e
\(e_{t_1}\) and \(e_{t_2}\)) at a loss function \(g(\bullet)\). The loss
function can be choosen flexibly as long as it is zero when the error is
zero, is strictly positive for all values on the real line and, it
increases with an increase in the absolute value of the error size. The
loss differential between two forecasts at time \(t\) is then defined
by, \[ d_t =  g(e_{t1}) - g(e_{t2})\] for \(t = (1, 2, ..., M)\) where
\(M\) is the total number of forecasts that have been generated from
each model. We then test whether, \[ \mu = E[d_t] = 0 \] by conducting
the following hypothesis test;\\
Null hypothesis: \[ E[d_t] = 0  \,\,\,\, \forall t\]\\
Alternative hypothesis: \[ E[d_t] \neq 0 \]\\
Test statistic:
\[DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} \]
